Certified Knowledge Compilation \\ with Application to Verified Model Counting \\ Experimental Details

### 2023-11-16 doc/JAIR/jair-data.tex
Work on optimization data

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.70]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.30,0.96)}},
      legend cell align={left},
                              x post scale=2.0, y post scale=2.0,
                              xmode=log,xmin=1e2,xmax=3e7,
                              xtick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000, 1e10}, 
                              xticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$, $10^{10}$},
                              ymode=log,ymin=1e2,ymax=3e7,
                              ytick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000, 1e10}, 
                              yticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$, $10^{10}$},
                              xlabel={Monolithic proof (clauses)}, ylabel={Structural proof + check (clauses)},
            ]

@@

\section*{Computing Counts and Weighted Counts}

\begin{itemize}
\item Implemented C package to represent numbers of form $a \cdot 2^b \cdot 5^c$.
  \begin{itemize}
  \item For unweighted counts, first compute density and then scale by $2^n$
  \item For unweighted counts, perform exact decimal arithmetic
  \item The set of numbers of this form is not closed under a reciprocal operation, but the weights in the competion are either of the form
    $w(x) + w(\overline{x}) = 1.0$ or $w(x) = w(\overline{x}) = 1$.  The latter case sums to 2, with reciprocal $1/2$.
  \end{itemize}
\item Can implement having $a$ be integer of arbitrary size and $b$ and $c$ being 32-bit integers.
  \begin{itemize}
    \item Represent $a$ by ``digits,'' each ranging in value from $0$ to $10^9-1$.
    \item Fits in 32-bit number
    \item Easy scaling by 10
    \item Easy conversion to/from decimal string
  \end{itemize}
\item Weighted counts ranged up to 260,909 decimal digits.
\end{itemize}

\section*{Time to Generate, Validate, and Count CPOG}

\begin{itemize}
\item Min 0.012, Average 1,329.8, Max 13,198.3, Median 75.4
\item Ratio to time for D4: Min 0.51, Harmonic Mean 5.87, Max 465.8
\end{itemize}

\input{figure-runtime-d4-verify}

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.90]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log, ymin=100, ymax=3000000000,
      legend style={at={(0.30,0.98)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=1.8,
                              xmode=log, xmin=0.01,xmax=4000,
                              xtick={0.01, 0.1,1.0,10,100,1000,10000}, xticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}, {10,000}},
                              ytick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000, 1e10}, yticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$,$10^{10}$},
                              xlabel={D4 runtime (seconds)}, ylabel={Defining Clauses},
%                              title={D4 Defining Clause Generation}
            ]
    \input{data-formatted/time+defining-success}
    \input{data-formatted/time+defining-onesided-only}
    \input{data-formatted/time+defining-failure}
    \legend{
      \scriptsize \textsf{Full validation},
      \scriptsize \textsf{One-sided validation},
      \scriptsize \textsf{Resource limit exceeded},
    }
 \end{axis}
\end{tikzpicture}
} % centering
\caption{Number of defining clauses in generated POG as function of D4 runtime}
\label{fig:generate:d4:log}
\end{figure}

\section*{Generated CPOG Files}

\begin{itemize}
\item Proof Clauses: Min 554, Median 1,719,245, Average 45,001,945, Max 770,382,773
\item Ratio to number of defining clauses: Min 1.54, Harmonic Mean 3.13, Max 6072.96
\end{itemize}

\input{figure-clauses-d4-result}

@@

\item[Literal grouping:] For a product node with multiple literal
  arguments, those that cannot be validated by BCP are combined as
  arguments to a newly defined product node.  The extension variable
  associated with this node is validated, and the proof is then used
  to validate the literals.  This reduces the number of
  calls to the SAT solver to at most one per product node.
\end{description}

>>>>

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.70]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.30,0.96)}},
      legend cell align={left},
                              x post scale=2.0, y post scale=2.0,
                              xmode=log,xmin=1e2,xmax=3e7,
                              xtick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000, 1e10}, 
                              xticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$, $10^{10}$},
                              ymode=log,ymin=1e2,ymax=3e7,
                              ytick={100,1000, 10000, 100000, 1000000, 10000000, 100000000, 1000000000, 1e10}, 
                              yticklabels={$10^2$,$10^3$,$10^4$,$10^5$,$10^6$,$10^7$,$10^8$,$10^9$, $10^{10}$},
                              xlabel={Monolithic proof (clauses)}, ylabel={Structural proof  (clauses)},
            ]

@@

@@

\item[Literal grouping:] For a product node with multiple literal
  arguments, those that cannot be validated by BCP are combined as
  arguments to a newly defined product node.  The extension variable
  associated with this node is validated, and the proof is then used
  to validate the literals.  This reduces the number of
  calls to the SAT solver to at most one per product node.
\end{description}
These optimizations are only applicable when the forward implication
proof is generated by the structural method, and so we concentrate on
the 50 benchmarks (from the smaller set of 90 benchmarks having at
most $10^7$ defining clauses) having tree ratios greater than $5.0$.
